Gametic Selection, Meiotic Drive, Sex Ratio Bias, and Transitions Between sex-determination systems

### 2017-07-06 LaTeX/Turnover.tex
incorporating mscotts comments

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
First, Fisherian sex-ratio selection will favour the spread of a neo-Y if the ancestral zygotic sex ratio is biased towards females, $\zeta>1/2$, and vice versa for a neo-W (i.e., $\zeta>1/2$ causes the first factor of the $\lambda_{mi}$ to be greater than one for a neo-Y).
However, the spread of a neo-Y (neo-W) also depends on the male (female) fitness of alleles that they are associated with. %, see terms in curly brackets. 
Second, invasion by a neo-Y (neo-W) allele does not directly depend on the fitness of female (male) diploids (although they indirectly affect invasion by determining the allele frequencies $p^\female_X$ and $\hat{p}^\male$).
%involve any female (male) diploid selection terms ($w_{g}^{\female}$)
This is because a dominant neo-Y (neo-W) is always found in males (females), and therefore the frequency of the neo-Y (neo-W) allele, $m$, only changes in males (females). 
%Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
%Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids. 
Finally, invasions by a neo-Y and a neo-W are qualitatively different.
This is because a gamete containing a neo-Y always pairs with the same gamete type as a male gamete without the neo-Y does (both pair with a female gamete containing an X), and both develop into males.
Meanwhile a gamete with a neo-W can pair with an X or Y male gamete, and develop into a female, while female gametes without the neo-W must pair with a male gamete containing an X to remain female.
This is consequential because it means that females with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating.
\textcolor{blue}{nice}
%the diploid fitness terms in Table \ref{tab:haplotype_growth} are weighted by the probability of producing those genotypes through matings with gametes of the opposite sex. 
%For example, matings between a neo-Y-bearing male gamete and an $A$-bearing female gamete occur with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$. 
%This is the same probability that an ancestral-Y-bearing male gamete will inherit an $A$ allele via mating with a female gamete. 
%On the other hand, 
%The probability that a neo-W bearing female gamete mates with an $A$-bearing male gamete is $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the frequency of the A allele among both X- and Y-bearing male gametes. 
%That is, in the case of a neo-W, female diploids can result from matings with either an X-bearing or a Y-bearing sperm, resulting in zygotes that will develop as females. 
%However, females that do not carry the neo-W only result from matings with X-bearing sperm. 
%Therefore, eggs with and without a neo-W can differ in the frequency of $A$ alleles they obtain from matings with male gametes. 
%Invasions by a neo-Y and a neo-W differ in this respect because sperm with or without a neo-Y allele both mate with X-bearing female gametes only.

@@

A neo-Y will never invade an ancestral XY system that already has tight linkage with the locus under selection ($r=0$). 
%When then neo-Y is also tightly linked ($R=0$) a neo-Y will either remain linked to the $A$ allele or to the $a$ allele and so invasion is given directly by the larger of $\lambda_{mA}$ or $\lambda_{ma}$.
A neo-Y haplotype with the same allele as the ancestral Y is neutral ($\lambda_{ma}=1$) and does not change in frequency.
The other neo-Y haplotype will not spread ($\lambda_{mA}<1$) given that the initial equilibrium is stable. 
Therefore, a neo-Y mutation cannot spread ($\lambda \leq 1$) in an ancestral XY system that is at equilibrium with all selected loci within the non-recombining region around the SDR.
\textcolor{blue}{
In essence, through tight linkage with the \textbf{A} locus, the ancestral Y becomes strongly specialized on the allele that has the highest fitness across male haploid and diploid phases. 
Given that the ancestral Y is at this equilibrium, it is not possible for a neo-Y to create males that have higher fitness than the ancestral Y. 
}

\textcolor{blue}{I might be tempted to focus on the diploid selection only case here because this is a new/surprising result and we have less cool stuff to say about haploid selection beyond describing it affecting stuff (right?). I then might also leave the complication of whether there is a polymorphism on the X or not to the appendix??? Attempt 1:}

\textcolor{blue}{
%Briefly, the neo-W invades under the ancestral XY sex-determining system, the X chromosome may not be able to as specialised on a female-beneficial allele
%The full characteristic polynomials are given in the appendix (equations \ref{Binvasion} and \ref{Ainvasion}).
Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (the full invasion conditions are given in the appendix). 
Significantly, we note that it is possible for both neo-W haplotypes to spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Binvasion} and \ref{Ainvasion}), in which case neo-W invasion can occur regardless of linkage to the selected locus.
That is, selection on loci within the non-recombining region of the SDR can favour an unlinked neo-W. 
Although haploid selection can favour neo-W alleles because the ancestral sex ratio becomes male biased, this is not the only circumstance in which less tightly linked neo-W alleles invade. 
For example, unlinked neo-W alleles can invade in the absence of any haploid selection. 
This result is unexpected given the results of \cite{vanDoorn:2010hu}, who did not explicitly calculate equilibrium allele frequencies under tight linkage and generally concluded that heterogametic transitions occur when neo-sex-determining alleles are in tighter linkage with loci under sex-specific diploid selection. 
Therefore, we focus on cases where there is no haploid selection and discuss the effects of haploid selection in the appendix. 
}

\textcolor{blue}{
If we categorise the $a$ allele as being ancestrally `male-beneficial' via the fact that it is fixed on the Y, then $\lambda_{mA}>1$ indicates that the neo-W spreads when found with the ancestrally `female-beneficial' allele. 
Intuitively, this is possible because the ancestral X chromosome is not able to specialise on the `female-beneficial' allele due to the fact that X's are sometimes found in males. 
For example, a polymorphism of $A$ and $a$ alleles can be maintained on the X despite directional selection in favour of the $A$ allele in females ($s^\female>0$, $0<h^\female<1$) because the $a$ allele is favoured in males. 
W$A$ haplotypes will only create females with high fitness $AA$ or $Aa$ genotypes and can therefore have higher fitness than ancestral females, which sometimes also produce $aa$ females. 
Thus, the neo-W can spread by allowing increased specialization on female beneficial alleles.
%As another example, $AA$ individuals can have low fitness in females and yet the $A$ can be fixed on the X due to strong overdominance in males favouring X$A$Y$a$ males. 
%In this case, W$A$ haplotypes produce fewer unfit $AA$ females because they produce some females through matings with Y$a$-bearing sperm. 
}

\textcolor{blue}{
Given that the $a$ allele can be considered ancestrally `male-beneficial' because it is fixed on the Y, it might be surprising that neo-W$a$ haplotypes can be favoured by selection in females, i.e., $\lambda_{ma}>1$. 
Again, this occurs because ancestral X's sometimes experience selection in males, in which they will always be paired with a Y$a$. 
Hence, if there is overdominance in males, X$A$Y$a$ males have high fitness and the $A$ allele is favoured by selection on the X in males. 
Therefore, the X can be polymorphic or fixed for the $A$ allele despite favouring the $a$ allele during selection in females \textcolor{red}{(see also Lloyd 197?, Otto 2014)}. 
In such cases, neo-W$a$ haplotypes can spread. 
Indeed, it is possible for both W$A$ and W$a$ haplotypes to spread, as is the case when $AA$ individuals have low fitness in females yet the $A$ is fixed on the X due to strong overdominance in males.
Both neo-W$A$ and neo-W$a$ haplotypes then produce fewer unfit $AA$ females.
This is true for the neo-W-$A$ haplotype because it can pair with a $Y-a$ haplotype and still be female. 
}

\textcolor{blue}{
We plot parameters under which $\lambda_{mA}>1$ and $\lambda_{ma}>1$ in figure XXX with diploid selection only. 
Wherever both haplotypes have positive growth rates, invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where at least one of these haplotypic growth rates is positive, a new neo-W feminizing mutation as long as \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is low enough and yet still indicates that more loosely linked sex-determining regions can spread. 
}

Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (given in detail in the appendix). 
%The full characteristic polynomials are given in the appendix (equations \ref{Binvasion} and \ref{Ainvasion}).
Briefly, neo-W-$A$ and/or neo-W-$a$ haplotypes can spread when rare in the absence of recombination ($\lambda_{ma}>1$ and/or $\lambda_{mA}>1$), depending on the ancestral sex-ratio and allele frequencies. 
Below we discuss the main forces determining the spread of these neo-W haplotypes and the impact of recombination for the overall success of the neo-W.
To simplify our discussion we first outline the potential effects of haploid selection and then consider diploid selection in its absence.

\textcolor{blue}{This explanation is clearer than the one I put in the appendix, if we drop it here, maybe we should try and re-word the corresponding part of the appendix using this paragraph}
Haploid selection impacts the spread of neo-W haplotypes in three ways.
Firstly, the zygotic sex ratio becomes male biased ($\zeta<1/2$) when the $a$ allele (which is fixed on the Y) is favoured during competition among male gametes or by meiotic drive in males. 
This facilitates the spread of a neo-W because neo-W alleles cause the zygotes that carry them to develop as the rarer, female, sex. 
Secondly, haploid selection in males affects the diploid genotypes of females by altering the allele frequencies in the male gametes that female gametes pair with.
For instance, because an epistatically dominant neo-W always causes its carrier to become female, it creates females who carry either $Y-a$ or $X$ genotypes from their father.
Thus, because when there is a polymorphism the $X$ carries some non-zero frequency of $A$, haploid selection in males impacts the diploid genotypes of females (e.g., creating more $Aa$ females when drive in males favours $Y-a$).
How this affects the spread of the neo-W then depends on diploid and haploid selection in females.
Thirdly, female drive and gamete competition directly select on neo-W haplotypes.
Drive for $A$ in females favours neo-W-$A$ haplotypes, at a cost to neo-W-$a$ haplotypes, and vice-versa when there is drive for $a$.
The impact of this drive depends on how often XX and neo-W females are heterozygous.
Competition among female gametes acts similarly, and depends on the frequency of $A$ on resident X chromosomes (e.g., competition among eggs has no affect on the initial spread of the neo-W-$A$ haplotype when $A$ is fixed on the X).
Because haploid selection in females favours one neo-W haplotype at the expense of the other, recombination off the favoured background becomes more detrimental as it becomes more favoured. Thus higher rates of recombination between the neo-W and the selected locus, $R$, can lead to smaller leading eigenvalues when there is haploid selection in females.

In the absence of haploid selection and with the $A$ allele is fixed on the X, it is possible for both neo-W haplotypes can spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Binvasion}), and thus neo-W invasion can occur regardless of its linkage to the selected locus.
Invasion does not occur with purely sexually-antagonistic selection (i.e., $a$ directionally favoured in males and $A$ directionally favoured in females) because the X is then already as specialized as possible on the female sex.
However, if, for example, $AA$ individuals suffer a fitness cost in females, yet $A$ is fixed on the X due to strong overdominance in males, both neo-W-$A$ and neo-W-$a$ haplotypes spread because they produce fewer unfit $AA$ females and never experience counter-selection in males.
This is true even for the neo-W-$A$ haplotype because it can pair with a $Y-a$ haplotype and still be female. 
When both haplotypes can spread alone the rate of recombination between the neo-W and the selected locus, $R$, does cannot prevent invasion, and thus the system can evolve looser sex-linkage (e.g., the neo-W could arise on an autosome, $R=1/2$).
Even when only one haplotype can spread, invasion can still occur up to some positive rate of recombination, $R>0$ (as long as equation \ref{eq:lambdasGen} is satisfied). 
That looser sex-linkage can evolve is contrary to the conclusions of \cite{vanDoorn:2010hu}, who did not explicitly calculate invasion fitness under ancestrally tight sex-linkage. \textcolor{blue}{or, just didn't calculate the equilibria (A) an (B).}
Similar scenarios have been shown to select for a modifier that increases recombination between the sexes \citep[green regions of Figure 2 in][]{Otto2014}.
\textcolor{blue}{n.b., the recombination regions are only subsets mostly because you also generate a really unfit Y haplotype through recombination. Probably not worth mentioning.}

In the absence of haploid selection it is also possible for a neo-W to invade when there is a stable polymorphism at the $\mathbf{A}$ locus on X chromosomes.
For example, overdominance in males and strong directional selection for $a$ in females creates a scenario that favours the spread of both neo-W haplotypes at equilibrium ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Ainvasion}), as both haplotypes bring more $a$ alleles into females and never experience counter-selection in males.
Thus, as in the case of the $A$ being fixed on the X, looser sex-linkage can evolve with a polymorphic X (i.e., $\lambda>1$ with $R>0$) and this is expected under the same scenarios that select for a modifier that increases recombination between the sex chromosomes \citep[blue regions of Figure 2 in][]{Otto2014}.

@@

The neo-sex-determining allele, $m$, will spread if $\lambda_{m,XY}>1$. 
Equation \eqref{eq:lambda_neoY} demonstrates that under weak selection a neo-Y will invade if and only if it is more closely linked to the selected locus than the ancestral sex-determining region (i.e., if $R<r$, note that $V_{A}$ and $S_{A}^2$ are strictly positive as long as \textbf{A} is polymorphic). 
This echoes our tight linkage results above and the results of \citet{vanDoorn:2007eu}, who considered diploid selection only and also found that homogametic transitions (XY to XY or ZW to ZW) can occur when the neo-sex-determining locus is more closely linked to a locus under sexually-antagonistic selection. 
%Because tighter linkage between the sex-determining region and a selected locus evolves during these transitions, stronger associations between selected alleles and diploid sexes can build up by selection (e.g., between male-beneficial alleles and the male-determining allele, Y).%, increasing diploid mean fitness (Figure \ref{fig:Combination_MeanFit}).

Equation \eqref{eq:lambda_neoW} shows that if there is no haploid selection ($t^\Hermaphrodite=\alpha^\Hermaphrodite_{\Delta}=0$), as considered by \citet{vanDoorn:2010hu}, with weak selection the spread of a neo-W is equivalent to the spread of a neo-Y ($\lambda_{W',XY}=\lambda_{Y',XY}$), such that heterogametic transitions (XY to ZW or ZW to XY) can also occur only if the neo-sex-determining region is more closely linked to a locus under selection ($R<r$). 
However, if there is any haploid selection, the additional term in equation \eqref{eq:lambda_neoW} can be positive, which can allow, for example, neo-W invasion ($\lambda_{W',XY}>1$) even when the neo-sex-determining region is less closely linked to the selected locus ($R>r$). 
These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
Mean fitness can therefore decrease with a heterogametic transition (Figure \ref{fig:Combination_MeanFit}B,D).

@@

@@

>>>>

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
First, Fisherian sex-ratio selection will favour the spread of a neo-Y if the ancestral zygotic sex ratio is biased towards females, $\zeta>1/2$, and vice versa for a neo-W (i.e., $\zeta>1/2$ causes the first factor of the $\lambda_{mi}$ to be greater than one for a neo-Y).
However, the spread of a neo-Y (neo-W) also depends on the male (female) fitness of alleles that they are associated with. %, see terms in curly brackets. 
Second, invasion by a neo-Y (neo-W) allele does not directly depend on the fitness of female (male) diploids (although they indirectly affect invasion by determining the allele frequencies $p^\female_X$ and $\hat{p}^\male$).
%involve any female (male) diploid selection terms ($w_{g}^{\female}$)
This is because a dominant neo-Y (neo-W) is always found in males (females), and therefore the frequency of the neo-Y (neo-W) allele, $m$, only changes in males (females). 
%Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
%Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids. 
Finally, invasions by a neo-Y and a neo-W are qualitatively different.
This is because a gamete containing a neo-Y always pairs with the same gamete type as a male gamete without the neo-Y does (both pair with a female gamete containing an X), and both develop into males.
Meanwhile a gamete with a neo-W can pair with an X or Y male gamete, and develop into a female, while female gametes without the neo-W must pair with a male gamete containing an X to remain female.
This is consequential because it means that females with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating.
%\textcolor{blue}{nice}
%the diploid fitness terms in Table \ref{tab:haplotype_growth} are weighted by the probability of producing those genotypes through matings with gametes of the opposite sex. 
%For example, matings between a neo-Y-bearing male gamete and an $A$-bearing female gamete occur with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$. 
%This is the same probability that an ancestral-Y-bearing male gamete will inherit an $A$ allele via mating with a female gamete. 
%On the other hand, 
%The probability that a neo-W bearing female gamete mates with an $A$-bearing male gamete is $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the frequency of the A allele among both X- and Y-bearing male gametes. 
%That is, in the case of a neo-W, female diploids can result from matings with either an X-bearing or a Y-bearing sperm, resulting in zygotes that will develop as females. 
%However, females that do not carry the neo-W only result from matings with X-bearing sperm. 
%Therefore, eggs with and without a neo-W can differ in the frequency of $A$ alleles they obtain from matings with male gametes. 
%Invasions by a neo-Y and a neo-W differ in this respect because sperm with or without a neo-Y allele both mate with X-bearing female gametes only.

@@

A neo-Y will never invade an ancestral XY system that already has tight linkage with the locus under selection ($r=0$). 
%When then neo-Y is also tightly linked ($R=0$) a neo-Y will either remain linked to the $A$ allele or to the $a$ allele and so invasion is given directly by the larger of $\lambda_{mA}$ or $\lambda_{ma}$.
A neo-Y haplotype with the same allele as the ancestral Y is neutral ($\lambda_{ma}=1$) and does not change in frequency.
The other neo-Y haplotype will not spread ($\lambda_{mA}<1$) given that the initial equilibrium is stable. 
Therefore, a neo-Y mutation cannot spread ($\lambda \leq 1$) in an ancestral XY system that is at equilibrium with all selected loci within the non-recombining region around the SDR.
In essence, through tight linkage with the \textbf{A} locus, the ancestral Y becomes strongly specialized on the allele that has the highest fitness across male haploid and diploid phases. 
Given that the ancestral Y is at this equilibrium, it is not possible for a neo-Y to create males that have higher fitness than the ancestral Y.

%\textcolor{blue}{I might be tempted to focus on the diploid selection only case here because this is a new/surprising result and we have less cool stuff to say about haploid selection beyond describing it affecting stuff (right?). I then might also leave the complication of whether there is a polymorphism on the X or not to the appendix??? Attempt 1:}

%\textcolor{blue}{
%Briefly, the neo-W invades under the ancestral XY sex-determining system, the X chromosome may not be able to as specialised on a female-beneficial allele
%The full characteristic polynomials are given in the appendix (equations \ref{Binvasion} and \ref{Ainvasion}).
Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (the full invasion conditions are given in the appendix; equations \ref{Binvasion} and \ref{Ainvasion}). 
Significantly, we note that it is possible for both neo-W haplotypes to spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$), in which case neo-W invasion can occur regardless of linkage to the selected locus.
That is, selection on loci within the non-recombining region of the SDR can favour the invasion of a less closely linked neo-W (e.g., on an autosome). 
Although haploid selection can favour neo-W alleles because the ancestral sex ratio becomes male biased, this is not the only circumstance in which less tightly linked neo-W alleles invade. 
For example, unlinked neo-W alleles can invade in the absence of any haploid selection. 
This result is unexpected given the results of \cite{vanDoorn:2010hu}, who did not explicitly calculate equilibrium allele frequencies under tight linkage and generally concluded that heterogametic transitions occur when neo-sex-determining alleles are in tighter linkage with loci under sex-specific diploid selection. 
Therefore, we focus on cases where there is no haploid selection and discuss the effects of haploid selection in the appendix. 
%}

%\textcolor{blue}{
If we categorise the $a$ allele as being ancestrally `male-beneficial' via the fact that it is fixed on the Y, then $\lambda_{mA}>1$ indicates that the neo-W spreads when found with the ancestrally `female-beneficial' allele. 
Intuitively, this is possible because the ancestral X chromosome is not able to specialise on the `female-beneficial' allele due to the fact that X's are sometimes found in males. 
For example, a polymorphism of $A$ and $a$ alleles can be maintained on the X despite directional selection in favour of the $A$ allele in females ($s^\female>0$, $0<h^\female<1$) because the $a$ allele is favoured in males. 
W-$A$ haplotypes will only create females with high fitness $AA$ or $Aa$ genotypes and can therefore have higher fitness than ancestral females, which sometimes also produce $aa$ females. 
Thus, the neo-W can spread by allowing increased specialization on female beneficial alleles.
%As another example, $AA$ individuals can have low fitness in females and yet the $A$ can be fixed on the X due to strong overdominance in males favouring X$A$Y$a$ males. 
%In this case, W$A$ haplotypes produce fewer unfit $AA$ females because they produce some females through matings with Y$a$-bearing sperm. 
%}

%\textcolor{blue}{
Given that the $a$ allele can be considered ancestrally `male-beneficial' because it is fixed on the Y, it might be surprising that neo-W-$a$ haplotypes can be favoured by selection in females ($\lambda_{ma}>1$). 
Again, this occurs because ancestral X's also experience selection in males, in which they will always be paired with a Y-$a$. 
Hence, if there is overdominance in males, X-$A$ Y-$a$ males have high fitness and the $A$ allele is favoured by selection on the X in males. 
Therefore, the X can be polymorphic or even fixed for the $A$ allele despite favouring the $a$ allele during selection in females \citep{Lloyd1977,Otto2014}. 
In such cases, neo-W-$a$ haplotypes, which are never found in males, can spread because they both create more $Aa$ and $aa$ females when pairing with an X from males and they bring Y-$a$ haplotypes into females, in which case females are always $aa$. 
Indeed, it is possible for both W-$A$ and W-$a$ haplotypes to spread, as is the case when $AA$ individuals have low fitness in females yet the $A$ is fixed on the X due to strong overdominance in males.
Both neo-W-$A$ and neo-W-$a$ haplotypes then produce fewer unfit $AA$ females.
This is true for the neo-W-$A$ haplotype because it can pair with a $Y-a$ haplotype and still be female. 
%}

%\textcolor{blue}{
In Figure \ref{fig:regionplot_ 2a} we show a region of parameter space under which both neo-W haplotypes invade ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) when there is overdominance in females and no haploid selection \citep[corresponding to Figure 2a in][]{Otto2014}. 
Wherever both haplotypes have positive growth rates (gray region of Figure \ref{fig:regionplot_ 2a}), invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where only one haplotype can spread (white region of Figure \ref{fig:regionplot_ 2a}), a neo-W can invade as long as equation \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is small enough and yet still indicates that more loosely linked sex-determining regions can spread.
%}

%Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (given in detail in the appendix). 
%%The full characteristic polynomials are given in the appendix (equations \ref{Binvasion} and \ref{Ainvasion}).
%Briefly, neo-W-$A$ and/or neo-W-$a$ haplotypes can spread when rare in the absence of recombination ($\lambda_{ma}>1$ and/or $\lambda_{mA}>1$), depending on the ancestral sex-ratio and allele frequencies. 
%Below we discuss the main forces determining the spread of these neo-W haplotypes and the impact of recombination for the overall success of the neo-W.
%To simplify our discussion we first outline the potential effects of haploid selection and then consider diploid selection in its absence.

%\textcolor{blue}{This explanation is clearer than the one I put in the appendix, if we drop it here, maybe we should try and re-word the corresponding part of the appendix using this paragraph}
%Haploid selection impacts the spread of neo-W haplotypes in three ways.
%Firstly, the zygotic sex ratio becomes male biased ($\zeta<1/2$) when the $a$ allele (which is fixed on the Y) is favoured during competition among male gametes or by meiotic drive in males. 
%This facilitates the spread of a neo-W because neo-W alleles cause the zygotes that carry them to develop as the rarer, female, sex. 
%Secondly, haploid selection in males affects the diploid genotypes of females by altering the allele frequencies in the male gametes that female gametes pair with.
%For instance, because an epistatically dominant neo-W always causes its carrier to become female, it creates females who carry either $Y-a$ or $X$ genotypes from their father.
%Thus, because when there is a polymorphism the $X$ carries some non-zero frequency of $A$, haploid selection in males impacts the diploid genotypes of females (e.g., creating more $Aa$ females when drive in males favours $Y-a$).
%How this affects the spread of the neo-W then depends on diploid and haploid selection in females.
%Thirdly, female drive and gamete competition directly select on neo-W haplotypes.
%Drive for $A$ in females favours neo-W-$A$ haplotypes, at a cost to neo-W-$a$ haplotypes, and vice-versa when there is drive for $a$.
%The impact of this drive depends on how often XX and neo-W females are heterozygous.
%Competition among female gametes acts similarly, and depends on the frequency of $A$ on resident X chromosomes (e.g., competition among eggs has no affect on the initial spread of the neo-W-$A$ haplotype when $A$ is fixed on the X).
%Because haploid selection in females favours one neo-W haplotype at the expense of the other, recombination off the favoured background becomes more detrimental as it becomes more favoured. Thus higher rates of recombination between the neo-W and the selected locus, $R$, can lead to smaller leading eigenvalues when there is haploid selection in females.

%In the absence of haploid selection and with the $A$ allele is fixed on the X, it is possible for both neo-W haplotypes can spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Binvasion}), and thus neo-W invasion can occur regardless of its linkage to the selected locus.
%Invasion does not occur with purely sexually-antagonistic selection (i.e., $a$ directionally favoured in males and $A$ directionally favoured in females) because the X is then already as specialized as possible on the female sex.
%However, if, for example, $AA$ individuals suffer a fitness cost in females, yet $A$ is fixed on the X due to strong overdominance in males, both neo-W-$A$ and neo-W-$a$ haplotypes spread because they produce fewer unfit $AA$ females and never experience counter-selection in males.
%This is true even for the neo-W-$A$ haplotype because it can pair with a $Y-a$ haplotype and still be female. 
%When both haplotypes can spread alone the rate of recombination between the neo-W and the selected locus, $R$, does cannot prevent invasion, and thus the system can evolve looser sex-linkage (e.g., the neo-W could arise on an autosome, $R=1/2$).
%Even when only one haplotype can spread, invasion can still occur up to some positive rate of recombination, $R>0$ (as long as equation \ref{eq:lambdasGen} is satisfied). 
%That looser sex-linkage can evolve is contrary to the conclusions of \cite{vanDoorn:2010hu}, who did not explicitly calculate invasion fitness under ancestrally tight sex-linkage. \textcolor{blue}{or, just didn't calculate the equilibria (A) an (B).}
%Similar scenarios have been shown to select for a modifier that increases recombination between the sexes \citep[green regions of Figure 2 in][]{Otto2014}.
%\textcolor{blue}{n.b., the recombination regions are only subsets mostly because you also generate a really unfit Y haplotype through recombination. Probably not worth mentioning.}
%
%In the absence of haploid selection it is also possible for a neo-W to invade when there is a stable polymorphism at the $\mathbf{A}$ locus on X chromosomes.
%For example, overdominance in males and strong directional selection for $a$ in females creates a scenario that favours the spread of both neo-W haplotypes at equilibrium ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Ainvasion}), as both haplotypes bring more $a$ alleles into females and never experience counter-selection in males.
%Thus, as in the case of the $A$ being fixed on the X, looser sex-linkage can evolve with a polymorphic X (i.e., $\lambda>1$ with $R>0$) and this is expected under the same scenarios that select for a modifier that increases recombination between the sex chromosomes \citep[blue regions of Figure 2 in][]{Otto2014}.

@@

The neo-sex-determining allele, $m$, will spread if $\lambda_{m,XY}>1$. 
Equation \eqref{eq:lambda_neoY} demonstrates that under weak selection a neo-Y will invade if and only if it is more closely linked to the selected locus than the ancestral sex-determining region (i.e., if $R<r$; note that $V_{A}$ and $S_{A}^2$ are strictly positive as long as \textbf{A} is polymorphic). 
This echoes our tight linkage results above and the results of \citet{vanDoorn:2007eu}, who considered diploid selection only and also found that homogametic transitions (XY to XY or ZW to ZW) can occur when the neo-sex-determining locus is more closely linked to a locus under sexually-antagonistic selection. 
%Because tighter linkage between the sex-determining region and a selected locus evolves during these transitions, stronger associations between selected alleles and diploid sexes can build up by selection (e.g., between male-beneficial alleles and the male-determining allele, Y).%, increasing diploid mean fitness (Figure \ref{fig:Combination_MeanFit}).

Equation \eqref{eq:lambda_neoW} shows that, in contrast to the tight linkage results of the previous section, if there is no haploid selection ($t^\Hermaphrodite=\alpha^\Hermaphrodite_{\Delta}=0$), as considered by \citet{vanDoorn:2010hu} and weak selection, the spread of a neo-W is equivalent to the spread of a neo-Y ($\lambda_{W',XY}=\lambda_{Y',XY}$), such that heterogametic transitions (XY to ZW or ZW to XY) can also occur only if the neo-sex-determining region is more closely linked to a locus under selection ($R<r$). 
However, if there is any haploid selection, the additional term in equation \eqref{eq:lambda_neoW} can be positive, which can allow, for example, neo-W invasion ($\lambda_{W',XY}>1$) even when the neo-sex-determining region is less closely linked to the selected locus ($R>r$). 
These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
Mean diploid fitness therefore decreases during heterogametic transitions that create looser sex-linkage (Figure \ref{fig:Combination_MeanFit}B,D).

@@

Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
We will also disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
We might therefore predict a difference when we  XY to ZW transitions 
However, under weak selection we find that invasion by a homogametic neo-sex-determining allele (XY to XY or ZW to ZW) or by a heterogametic neo-sex-determining allele (XY to ZW or ZW to XY) occur under the same conditions. 
This is because, with weak selection, male biased zygotic sex ratios ($\zeta-1/2$) favouring the invasion of a neo-W are of the same order as meiotic drive favouring the invasion of a driving ne $\epsilon^2$.

Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Haplotype invasion region plot
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] () {
	\includegraphics[width=0.5\linewidth]{regionplot_2a}
	\includegraphics[width=0.5\linewidth]{regionplot_hap}
	};
\end{tikzpicture}
\caption{
Gray region depicts the relative fitnesses of male homozygotes, $w_AA^\male$ and $w_aa^\male$, that allow both neo-W haplotypes to invade, from any stable resident equilibria (equations \ref{eq:tightequil}).
In the white region neo-W haplotypes paired with the allele fixed on the Y cannot invade.
Parameters as in \cite{Otto2014} Figure 2a: $w_A^\female=w_a^\female=w_A^\male=w_a^\male=1$, $\alpha^\female=\alpha^\male=1/2$, $w_Aa^\female=w_Aa^\male=1$, and $w_AA^\female=w_aa^\female=0.75$, with $r=0$.
}
\label{fig:regionplot_ 2a}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
